An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors

### 2021-03-31 comment.tex
Updates from Overleaf

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=black,very thick}]
      \path [->] (A) edge node {$1-\sens$} (2,4.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
      \path [->] (F) edge node {$\spec$} (3.7,5);
      \path [-Bar] (D) edge (4.5, 2);
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of our model. A pool contains individuals
    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
    are infected). A negative pooled test ($\data=0$) occurs when
    three detection paths fail: A false-negative occurs for
    individuals $1$ and $2$, each with probability $(1-\sens)$.
    Additionally, no false-positive detection of SARS-CoV-2 occurs,
    with probability $\spec$. Individual $3$ is not infected and does
    not contribute to the probability of the pooled test result. The
    pooled test is positive ($\data=1$) if any path results in a
    detection.}\label{fig:likelihood}
\end{figure}

@@

\section*{Discussion}
%\subsection{Short recap}
Here, we presented a novel probabilistic model for the outcomes of
pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
assumption, we assume, based on data (see section XX), that multiple
infected individuals increase the likelihood of a positive pooled test
result. A direct consequence of our model is that false-negative rates
depend on infection prevalence. Specifically, low values of infection
prevalence increase $\err{exact}$. These results remain qualitatively
similar under varying parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{fig1}).

Our results lead to a conflict: low infection prevalence leads to high
efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
increasing false-negative rates. This is a fact almost disregarded in
\cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
when infection prevalence is low, such circumstances are exactly those
leading to a substantial increase in false-negative rates. Such
behavior of pooling is not implied by the common assumption.

@@

%% One important consequence of the efficiency charts is that
%% practitioners order RT-PCR reagents, they should expect \emph{lower}
%% scheme efficiency, and thus order more reagents.

%\subsection{Limitations}
Lacking large amounts of experimental data, it is hard to assess the
validity of our model. The data in \cite{Salazar} does allow us to
reject the common assumption but not much more. Our model assumes as
little as as possible, given that the standard assumptions are refuted
and dilution effects are not taken into consideration.

%\subsection{Ways to reduce errors}
Although our application focused on Dorfman pooling, our analysis
applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify Donald as
infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two;
otherwise, everyone are declared negative. Then the splitting is
repeated until resulting pools are negative or individuals are tested
separately. With an initial pool size of 32, Donald will necessarily
have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
in a single test, for the scheme to correctly identify him as
infected. Compare this to the Dorfman scheme that requires a positive
test in a pool of size $N=8$, and an additional single positive test
to identify Donald as infected. The hierarchical pooling scheme of
\cite{Lion, Kim} will necessarily yield more false-negatives than
Dorfman pooling --- there are additional places for it to fail.

Thus, finding improved pooling schemes is imperative. An especially
important consideration in designing such schemes is explicitly taking
the intrinsic PCR error rates into account. We (ref), as well as
others (ref), have developed such models, although their
implementation in real-world settings has not yet commenced.

%\subsection{Conclusion}
We have no doubt that further study of the assumptions underlying
pooling is due. Furthermore, adapting common group testing algorithms
\cite{Kim} to novel error models should be further studied.

@@

%% To conclude, pooling is an important technique that can improve
%% testing throughput in a cost-effective manner. Nevertheless, care must
%% be given to pooling schemes' false-negative rates, especially under
%% low infection prevalence settings.

>>>>

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=black,very thick}]
      \path [->] (A) edge node {$1-\sens$} (2,4.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
      \path [->] (F) edge node {$\spec$} (3.7,5);
      %\path [-Bar] (D) edge (4.5, 2);
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of our model. A pool contains individuals
    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
    are infected). A negative pooled test ($\data=0$) occurs when
    three detection paths fail: A false-negative occurs for
    individuals $1$ and $2$, each with probability $(1-\sens)$.
    Additionally, no false-positive detection of SARS-CoV-2 occurs,
    with probability $\spec$. Individual $3$ is not infected and does
    not contribute to the probability of the pooled test result.}\label{fig:likelihood}
\end{figure}

@@

\section*{Discussion}
%\subsection{Short recap}
Here, we presented a novel probabilistic model for the outcomes of
pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
assumption, we assume, based on data \cite{Salazar}, that multiple
infected individuals increase the likelihood of a positive pooled test
result. A direct consequence of our model is that false-negative rates
depend on infection prevalence. Specifically, low values of infection
prevalence increase $\err{exact}$. These results remain qualitatively
similar under varying parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{fig1}).
Our results lead to a conflict: low infection prevalence leads to high
efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
increasing false-negative rates.

% This is a fact almost disregarded in
% \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
% when infection prevalence is low, such circumstances are exactly those
% leading to a substantial increase in false-negative rates. Such
% behavior of pooling is not implied by the common assumption.

@@

%\subsection{Limitations}
%NONE

%\subsection{Ways to reduce errors}
Although our application focused on Dorfman pooling, our analysis
applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify an infected individual as
such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two;
otherwise, everyone are declared negative. Then the splitting is
repeated until resulting pools are negative or individuals are tested
separately. With an initial pool size of 32, Donald will necessarily
have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
in a single test, for the scheme to correctly identify him as
infected. Compare this to the Dorfman scheme that requires a positive
test in a pool of size $N=8$, and an additional single positive test
to identify Donald as infected. The hierarchical pooling scheme of
\cite{Lion, Kim} will necessarily yield more false-negatives than
Dorfman pooling --- there are additional places for it to fail.

Thus, finding improved pooling schemes is imperative. An especially
important consideration in designing such schemes is explicitly taking
the intrinsic PCR error rates into account. We \cite{DOPE}, as well as
others \cite{BayesianDorfman}, have developed such models, although their
implementation in real-world settings has not yet commenced.

%\subsection{Conclusion}

@@

%% Conclusion
To conclude, pooling is an important technique that can improve
testing throughput in a cost-effective manner. Nevertheless, care must
be given to pooling schemes' false-negative rates, especially under
low infection prevalence settings.
